The role of cilia-mediated cerebrospinal fluid flow in intraventricular transport in zebrafish embryos

### 2025-02-03 paper/main.tex
Begin working on Results...

Wholemount~\emph{in vivo} live imaging and photoconversion experiments were performed with zebrafish larvae at the 2 days post-fertilization (dpf) stage obtained from inbreeding of heterozygous \emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xuas:Signal-Dendra2}) adult animals, as described by D’Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}. Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+}) or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding. The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

@@

\subsection*{Wholemount zebrafish \emph{in vivo} live time-lapse imaging and photoconversion of Dendra2}
To image 2 dpf larval zebrafish, the fish were anesthetized in 0.013\% MS222 in AFW and mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100), to obtain a lateral view of the brain ventricles. After the fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify. After solidification, AFW containing 0.013\% MS222 was added, and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss) and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat) at room temperature.

@@

    \item A \emph{no-cilia/cardiac-only} flow model including contributions from the cardiac pulstaions but no cilia forces. Here, we set the tangential traction to zero on the cilia boundary $\Gc$, while prescribing the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$.  % Model II
    %\item Model II: Only cardiac pulsations are modeled, no cilia forces.

@@

%%--------------RESULTS--------------%%
\section*{Results}\label{sec:results}
\subsection*{Cerebrospinal fluid flow in the brain ventricles is compartmentalized}
With cilia forces and cardiac forces represented in the flow model (model version 0), we observe a vortex-structured flow pattern in the anterior and middle ventricles~(\cref{fig:figure1}, panel \textbf{e}). This rotational motion shows that cilia-mediated flow is dominating in these regions, as the vortex structures are a consequence of the applied tangential stresses on the ciliated boundaries combined with the assumption of fluid incompressibility. In the posterior ventricle, we observe directional flow as the pressure-driven flow is dominating in this region. The cardiac pressure forces induce a pulsatile motion of the cerebrospinal fluid synchronized with the cardiac cycle, with velocity magnitudes of the pulsatile motion being approximately an order of magnitude smaller than the velocities related to the cilia-driven motion. Over a cardiac cycle, the maximum velocity magnitude in the ventricles was 27.7 $\mu$m/s, located dorsally in the posterior part of the middle ventricle. This maximum velocity was attained at the time instant coinciding with the peak of the cardiac pulsatile flow flowing caudally, as a result of both the tangential stress boundary condition and the pulsatile pressure-driven flow being aligned with the rostrocaudal axis (the $x$ axis). A calculation using the observed maximum velocity of 27.7 $\mu$m/s, and choosing the approximate height of the middle ventricle $110 \ \mu$m as a characteristic length scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes flow.

@@

\subsection*{Comparing experimental and computational results}
Transport of photoconverted proteins is quantified by the change in fluorescence intensity $\Delta F=(F(t)-F_0)/F_0$ over time~(\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{a}). The solid lines are the mean values of $\Delta F$ for the fish cohort, with shaded regions covering one standard deviation. The first region of interest (ROI 1) is the laser-targeted region. Qualitatively, the fluorescence intensity change in this region resembles intensity dynamics observed in experiments that studied the photoconversion properties of Dendra2 fluorescent proteins~\cite{Makarov2014Steady-stateDendra2}. The transport to the other regions can be separated into two groups: the fluorescence intensity change $\Delta F$ for ROIs 2--4 in the middle ventricle are all similar, whereas the more distal regions ROI 5 and 6 (anterior and posterior ventricle, respectively) exhibit a slower increase in $\Delta F$. In ROIs 2--4, the $\Delta F$ curves are concave, similar to ROI 1, while in ROIs 5 and 6 the curves are convex. Although the fluorescence intensity increases throughout all of the experiments, the maximum $\Delta F$ is not necessarily attained at the final time of the experiments. At the final time, the mean fluoresence intensities in regions 2--4 all exceed 0.60 (0.63, 0.71, 0.67, respectively) of the final value $F_0$ in ROI 1, wheras ROIs 5 and 6 reach mean values of 0.29 and 0.18, respectively.

@@

The brain ventricular geometry varies significantly across zebrafish individuals, both healthy and sick animals. Motivated by an interest in how these variations can impact solute transport in the brain ventricles, we study the impact on simulation results when modifying the computational geometry and mesh in four different scenarios. Because our computational mesh is inflated by the dye-injection performed before image segmentation, we study four different mesh versions where we shrink: the fore-mid brain connection; the middle ventricle; the mid-hind brain connection; and all of the three above simultaenously. We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.

@@

\lyng{HH: Quantify differences in terms of percentage changes in the final mean concentration values, maybe with a bar plot?}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{Caption. \mk{MK: How do the different geoemtries compare in terms of volume, areas where cilia are active?} \lyng{HH: Good question, will add this information.}}
    \label{fig:figure6_sim_results_compare_geometries}
\end{figure}

>>>>

Whole-mount~\emph{in vivo} live imaging and photoconversion experiments were performed with zebrafish larvae at the 2 days post-fertilization (dpf) stage obtained from inbreeding of heterozygous \emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xuas:Signal-Dendra2}) adult animals, as described by D’Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}. Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+}) or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding. The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

@@

\subsection*{Whole-mount zebrafish \emph{in vivo} live time-lapse imaging and photoconversion of Dendra2}
To image 2 dpf larval zebrafish, the fish were anesthetized in 0.013\% MS222 in AFW and mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100), to obtain a lateral view of the brain ventricles. After the fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify. After solidification, AFW containing 0.013\% MS222 was added, and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss) and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat) at room temperature.

@@

    \item A \emph{no-cilia/cardiac-only} flow model including contributions from the cardiac pulsations but no cilia forces. Here, we set the tangential traction to zero on the cilia boundary $\Gc$, while prescribing the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$.  % Model II
    %\item Model II: Only cardiac pulsations are modeled, no cilia forces.

@@

%%--------------RESULTS--------------%%
\section*{Results}\label{sec:results}

\subsection*{Cerebrospinal fluid flow in the brain ventricles is compartmentalized}

\mer{Why?} To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we consider
the simulated CSF velocity and pressure when driven by both cilia and
cardiac pulsations. \mer{MER: Stopped here...} With cilia forces and
cardiac forces represented in the flow model (model version 0), we
observe a vortex-structured flow pattern in the anterior and middle
ventricles~(\cref{fig:figure1}, panel \textbf{e}). This rotational
motion shows that cilia-mediated flow is dominating in these regions,
as the vortex structures are a consequence of the applied tangential
stresses on the ciliated boundaries combined with the assumption of
fluid incompressibility. In the posterior ventricle, we observe
directional flow as the pressure-driven flow is dominating in this
region. The cardiac pressure forces induce a pulsatile motion of the
cerebrospinal fluid synchronized with the cardiac cycle, with velocity
magnitudes of the pulsatile motion being approximately an order of
magnitude smaller than the velocities related to the cilia-driven
motion. Over a cardiac cycle, the maximum velocity magnitude in the
ventricles was 27.7 $\mu$m/s, located dorsally in the posterior part
of the middle ventricle. This maximum velocity was attained at the
time instant coinciding with the peak of the cardiac pulsatile flow
flowing caudally, as a result of both the tangential stress boundary
condition and the pulsatile pressure-driven flow being aligned with
the rostrocaudal axis (the $x$ axis). A calculation using the observed
maximum velocity of 27.7 $\mu$m/s, and choosing the approximate height
of the middle ventricle $110 \ \mu$m as a characteristic length scale,
yields a Reynolds number $\mathrm{Re} \approx 0.004$, justifying the
assumption of Stokes flow.

@@

\subsection*{Comparing experimental and computational results}
Transport of photoconverted proteins is quantified by the change in fluorescence intensity $\Delta F=(F(t)-F_0)/F_0$ over time~(\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{a}). The solid lines are the mean values of $\Delta F$ for the fish cohort, with shaded regions covering one standard deviation. The first region of interest (ROI 1) is the laser-targeted region. Qualitatively, the fluorescence intensity change in this region resembles intensity dynamics observed in experiments that studied the photoconversion properties of Dendra2 fluorescent proteins~\cite{Makarov2014Steady-stateDendra2}. The transport to the other regions can be separated into two groups: the fluorescence intensity change $\Delta F$ for ROIs 2--4 in the middle ventricle are all similar, whereas the more distal regions ROI 5 and 6 (anterior and posterior ventricle, respectively) exhibit a slower increase in $\Delta F$. In ROIs 2--4, the $\Delta F$ curves are concave, similar to ROI 1, while in ROIs 5 and 6 the curves are convex. Although the fluorescence intensity increases throughout all of the experiments, the maximum $\Delta F$ is not necessarily attained at the final time of the experiments. At the final time, the mean fluorescence intensities in regions 2--4 all exceed 0.60 (0.63, 0.71, 0.67, respectively) of the final value $F_0$ in ROI 1, whereas ROIs 5 and 6 reach mean values of 0.29 and 0.18, respectively.

@@

The brain ventricular geometry varies significantly across zebrafish individuals, both healthy and sick animals. Motivated by an interest in how these variations can impact solute transport in the brain ventricles, we study the impact on simulation results when modifying the computational geometry and mesh in four different scenarios. Because our computational mesh is inflated by the dye-injection performed before image segmentation, we study four different mesh versions where we shrink: the fore-mid brain connection; the middle ventricle; the mid-hind brain connection; and all of the three above simultaneously. We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.

@@

\lyng{HH: Quantify differences in terms of percentage changes in the final mean concentration values, maybe with a bar plot?}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{Caption. \mk{MK: How do the different geometries compare in terms of volume, areas where cilia are active?} \lyng{HH: Good question, will add this information.}}
    \label{fig:figure6_sim_results_compare_geometries}
\end{figure}
